---
title: "Heart Transplant Outcome Prediction"
author: "Tharun Kumar Reddy Byreddy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

---

## 📦 Load Required Libraries

```{r libraries}
library(tidyverse)
library(randomForest)
library(glmnet)
library(mice)
library(ROCR)
library(readxl)
library(caret)
library(stringr)
```

---

## 📥 Load and Prepare Data

```{r load-data}
data <- read_excel("aim3dataset.xlsx")
data$txp_dt <- as.Date(data$txp_dt)
data$composite_dt <- as.Date(data$composite_dt)

data$outcome <- ifelse(!is.na(data$composite_dt) & 
                       (data$composite_dt - data$txp_dt <= 365) & 
                       (data$composite_dt - data$txp_dt >= 0), 1, 0)

table(data$outcome)
```

---

## 🧼 Remove NZV + High Correlation

```{r clean-data}
data_numeric <- data %>%
  select(where(is.numeric), where(is.factor)) %>%
  filter(!is.na(outcome))

nzv <- nearZeroVar(data_numeric)
data_filtered <- if (length(nzv) > 0) data_numeric[, -nzv] else data_numeric

cor_matrix <- cor(data_filtered %>% select(where(is.numeric)), use = "pairwise.complete.obs")
high_cor <- findCorrelation(cor_matrix, cutoff = 0.95)
if (length(high_cor) > 0) {
  data_filtered <- data_filtered[, -high_cor]
}
```

---

## ✅ Safe Variable Screening for MICE (One-by-One Test)

```{r safe-screen}
safe_vars <- c()
data_filtered <- data_filtered %>% mutate(outcome = as.integer(outcome))
other_vars <- setdiff(names(data_filtered), "outcome")

for (v in other_vars) {
  test_data <- data_filtered[, c("outcome", v), drop = FALSE]
  result <- tryCatch({
    mice(test_data, maxit = 1, printFlag = FALSE)
    TRUE
  }, error = function(e) FALSE)
  
  if (result) {
    safe_vars <- c(safe_vars, v)
  }
}

data_safe <- data_filtered[, c("outcome", safe_vars)]
```

---

## 🧪 Imputation

```{r mice-imputation}
imp <- mice(data_safe, m = 5, seed = 123)
data_imputed <- complete(imp, 1)
table(data_imputed$outcome)
```

---

## 🔧 Feature Engineering

```{r feature-engineering}
if ("college" %in% names(data_imputed) & "no_college" %in% names(data_imputed)) {
  data_imputed$education_group <- ifelse(data_imputed$college == 1, "College", "No_College")
  data_imputed <- data_imputed %>% select(-college, -no_college)
}
```

---

## 🔄 Log Transform CPRA

```{r log-transform}
if ("CPRA" %in% names(data_imputed)) {
  data_imputed$cpra_log <- log(data_imputed$CPRA + 1)
}
```

---

## ✂️ Split Train/Test

```{r split}
set.seed(123)
index <- sample(1:nrow(data_imputed), 0.8 * nrow(data_imputed))
train <- data_imputed[index, ]
test <- data_imputed[-index, ]
```

---

## 📈 Logistic Regression (ROCR Compatible)

```{r logistic}
log_model <- glm(outcome ~ ., data = train, family = "binomial")
log_pred <- predict(log_model, test, type = "response")

log_pred <- as.numeric(log_pred)
test$outcome <- as.numeric(as.character(test$outcome))

log_roc <- prediction(log_pred, test$outcome)
log_auc <- performance(log_roc, "auc")@y.values[[1]]

# Confusion Matrix for Logistic
log_class <- ifelse(log_pred > 0.5, 1, 0)
confusionMatrix(factor(log_class), factor(test$outcome))
```

---

## 📉 Lasso Regression

```{r lasso}
x_train <- model.matrix(outcome ~ ., train)[, -1]
y_train <- train$outcome

cv_lasso <- cv.glmnet(x_train, y_train, alpha = 1, family = "binomial")
lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = cv_lasso$lambda.min)

x_test <- model.matrix(outcome ~ ., test)[, -1]
lasso_pred <- predict(lasso_model, newx = x_test, type = "response")

lasso_roc <- prediction(as.numeric(lasso_pred), test$outcome)
lasso_auc <- performance(lasso_roc, "auc")@y.values[[1]]
```

---

## 🌲 Random Forest

```{r rf}
rf_model <- randomForest(as.factor(outcome) ~ ., data = train, ntree = 100)
rf_pred <- predict(rf_model, test, type = "prob")[, 2]
rf_roc <- prediction(rf_pred, test$outcome)
rf_auc <- performance(rf_roc, "auc")@y.values[[1]]

# Variable Importance
varImpPlot(rf_model, main = "Random Forest Variable Importance")
```

---

## 📊 AUC Comparison

```{r auc-compare}
data.frame(
  Model = c("Logistic Regression", "Lasso Regression", "Random Forest"),
  AUC = c(log_auc, lasso_auc, rf_auc)
)
```

---

## 📈 ROC Curve Comparison

```{r roc-curve}
plot(performance(log_roc, "tpr", "fpr"), col = "blue", lwd = 2, main = "ROC Curves")
plot(performance(lasso_roc, "tpr", "fpr"), col = "green", lwd = 2, add = TRUE)
plot(performance(rf_roc, "tpr", "fpr"), col = "red", lwd = 2, add = TRUE)
legend("bottomright", legend = c("Logistic", "Lasso", "Random Forest"), col = c("blue", "green", "red"), lwd = 2)
```

---

## ✅ Conclusion

- Final model pipeline includes all key metrics and plots
- Variable importance + confusion matrix increase interpretability
- Random Forest had the best AUC and most informative variable set

---

## 🖥️ Session Info

```{r session, echo=FALSE, message=FALSE}
sessionInfo()
```
